%%--------------------------------------------------
%% basics
% \documentclass[letterpaper,oneside,openany]{memoir}
\documentclass[ebook,10pt,oneside,openany,final]{memoir}
% \includeonly{declarations}

\usepackage[american]
           {babel}        % needed for iso dates
\usepackage[iso,american]
           {isodate}      % use iso format for dates
\usepackage[final]
           {listings}     % code listings
\usepackage{longtable}    % auto-breaking tables
\usepackage{ltcaption}    % fix captions for long tables
\usepackage{booktabs}     % fancy tables
\usepackage{relsize}      % provide relative font size changes
\usepackage{underscore}   % remove special status of '_' in ordinary text
\usepackage{verbatim}     % improved verbatim environment
\usepackage{parskip}      % handle non-indented paragraphs "properly"
\usepackage{array}        % new column definitions for tables
\usepackage[normalem]{ulem}
\usepackage{color}        % define colors for strikeouts and underlines
\usepackage{amsmath}      % additional math symbols
\usepackage{mathrsfs}     % mathscr font
\usepackage{multicol}
\usepackage{xspace}
\usepackage{fixme}
\usepackage{lmodern}
\usepackage{textcomp}     % provide \text{l,r}angle
\usepackage[T1]{fontenc}
\usepackage[pdftex, final]{graphicx}
\usepackage[pdftex,
            pdftitle={find\_last},
            pdfsubject={find\_last},
            pdfcreator={Zach Laine},
            bookmarks=true,
            bookmarksnumbered=true,
            pdfpagelabels=true,
            pdfpagemode=UseOutlines,
            pdfstartview=FitH,
            linktocpage=true,
            colorlinks=true,
            linkcolor=blue,
            plainpages=false
           ]{hyperref}
\usepackage{memhfixc}     % fix interactions between hyperref and memoir

\input{layout}
\input{styles}
\input{macros}
\input{tables}
\input{cxx}

%%--------------------------------------------------
%% fix interaction between hyperref and other
%% commands
\pdfstringdefDisableCommands{\def\smaller#1{#1}}
\pdfstringdefDisableCommands{\def\textbf#1{#1}}
\pdfstringdefDisableCommands{\def\raisebox#1{}}
\pdfstringdefDisableCommands{\def\hspace#1{}}

%%--------------------------------------------------
%% add special hyphenation rules
\hyphenation{tem-plate ex-am-ple in-put-it-er-a-tor name-space name-spaces}

\begin{document}
\chapterstyle{cppstd}
\pagestyle{cpppage}

%%--------------------------------------------------
%% configuration
\input{config}

%%--------------------------------------------------
%% front matter
\frontmatter
\include{front}

%%--------------------------------------------------
%% main body of the document
\mainmatter
\setglobalstyles

\include{revisions}

\section{Motivation and Scope}

Consider how finding the last element that is equal to `x` in a range is
typically done (for all the examples below, we assume a valid range of
elements \tcode{[first, last)}, and an iterator \tcode{it} within that range):

\begin{itemdecl}
    while (it-- != first) {
        if (*it == x) {
            // Use it here...
        }
    }
\end{itemdecl}

Raw loops are icky though.  Perhaps we should do a bit of extra work to allow
the use of \tcode{find()}:

\begin{itemdecl}
    auto rfirst = std::make_reverse_iterator(it);
    auto rlast = std::make_reverse_iterator(first);
    auto it = std::find(rfirst, rlast, x);
    // Use it here...
\end{itemdecl}

That seems nicer in that there is no raw loop, but it requires an unpleasant
amount of typing (and an associated lack of clarity).

Consider this instead:

\begin{itemdecl}
    auto it = std::find_last(first, it, x);
    // Use it here...
\end{itemdecl}

That's better!  It's a lot less verbose.

Let's consider for a moment the lack of clarity of the
\tcode{make_reverse_iterator()} code.  In a typical use of \tcode{find()}, I
search forward from the element I start from, including the element itself:

\begin{itemdecl}
    auto it = std::find(it, last, x); // Includes examination of *it.
\end{itemdecl}

However, using finding in reverse in the middle of a range leaves out the
element pointed to by the current iterator:

\begin{itemdecl}
    auto it = std::find( // Skips *it entirely.
        std::make_reverse_iterator(first),
        std::make_reverse_iterator(it),
        x);
\end{itemdecl}

That leads to code like this:

\begin{itemdecl}
    auto it = std::find( // Includes *it again!
        std::make_reverse_iterator(first),
        std::make_reverse_iterator(std::next(it)),
        x);
\end{itemdecl}

Though this looks like an off-by-one error, is is correct.  Moreover, even
though the use of \tcode{next()} is correct, it gets lost in noise of the rest
of the code, since it is so verbose.  Use \tcode{find_last()} makes things
clearer:

\begin{itemdecl}
    // Search, but don't include *it.
    auto it_1 = std::find_last(first, it, x);

    // Search, and include *it.
    auto it_2 = std::find_last(first, std::next(it), x);
\end{itemdecl}

The use of \tcode{next()} may at first appear like a mistake, until the reader
takes a moment to think things through.  In the \tcode{reverse_iterator}
version, this correctness is a lot harder to readily grasp.

\section{Design Considerations}

During LWG review it was noted that the original design does not account
for the principle of preserving useful information.  Specifically, with
sentinal-based ranges \tcode{find_last()} needs to calculate the end of the
range, but the api as specified does not return the end. At the time
\footnote{Dec 2020 LWG review notes https://wiki.edg.com/bin/view/Wg21fall2020/P1223-2020-12-11},
a small group had consensus to ask LEWG to change the return to include the end.

To incorporate the calculation of the end of the range would require changing the
signature from:

\begin{itemdecl}
    //current wording
    template<forward_iterator I, sentinel<I> S, class T, class Proj = identity>
      requires indirect_binary_predicate<ranges::equal_to, projected<I, Proj>, const T*>
      constexpr I find_last(I first, S last, const T& value, Proj proj = {});

    template<forward_range R, class T, class Proj = identity>
      requires indirect_binary_predicate<ranges::equal_to, projected<iterator_t<R>, Proj>, const T*>
      constexpr borrowed_iterator_t<R>
        find_last(R&& r, const T& value, Proj proj = {});
\end{itemdecl}

to:

\begin{itemdecl}
    // possible new wording
    template<forward_iterator I, sentinel<I> S, class T, class Proj = identity>
      requires indirect_binary_predicate<ranges::equal_to, projected<I, Proj>, const T*>
      constexpr subrange<I> find_last(I first, S last, const T& value, Proj proj = {});

    template<forward_range R, class T, class Proj = identity>
      requires indirect_binary_predicate<ranges::equal_to, projected<iterator_t<R>, Proj>, const T*>
      constexpr borrowed_subrange_t<R>
        find_last(R&& r, const T& value, Proj proj = {});
\end{itemdecl}

where the subrange would be \tcode{[i, last)} when found and \tcode{[last, last)}
when not found.

But there's a downside to this design.  The api is not inconsistent with the other
find algorithms \tcode{find and find_first}. Also, the usage in a typical use case
where only the found iterator is used the user has to discard the extra return
information. This was discussed at some length on the LEWG mailing list with only moderate support for the proposed change \footnote{LEWG list discussion 2021-12 http://lists.isocpp.org/lib-ext/2021/12/21613.php}.

Still, the author feels the change is an improvement and should be made.

SG9 discussed the paper on 2022-04-11 and was weakly in favor of the proposed change \footnote{SG9 results https://github.com/cplusplus/papers/issues/149#issuecomment-1097082737}.

Poll: We recommend making the change proposed in section 0.3 of P1223R3 for \tcode{find_last}, (making \tcode{find_last} returning a subrange) for C++23.
SF 	F 	N 	A 	SA
1 	4 	5 	1 	0

Outcome: Weak consensus in favor

LEWG mailing list review as of 2022-04-15 had 5 in favor and no against.

\section{Feature-Test Macro}

In addition to the wording that follows, add a new feature-test macro
\tcode{__cpp_lib_find_last}, with value \tcode{xxxxxx}.  This macro should also be
defined in \tcode{<algorithm>}.

\newcommand{\indexhdr}[1]{}
\newcommand{\indexlibrarymember}[2]{
}
\input{algorithms}

%%--------------------------------------------------
\include{acknowledgements}

%%--------------------------------------------------
%% End of document
\end{document}
